### 2023-04-04 manuscript/manuscript.qmd
Add bibliohgraphies

---
title: Short Paper
subtitle: A Short Subtitle
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliations: 
        - id: some-tech
          name: Some Institute of Technology
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Bob Security
    email: bob@example.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    note: |
      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  - name: Cat Memes
    email: cat@example.com
    affiliations:
        - ref: another-u
    note: Yet another author footnote.
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: [bibliography.bib, introduction.bib, man-qech_air-quality.bib]
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
editor: 
  markdown: 
    wrap: 72
---

# Introduction

Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
(QECH), or Queen's, as it's known locally, is the country's largest
public hospital. Designed in the Tropical Modernism style of
architecture popular within Africa in the late colonial period, Queen's
open-plan form sprawls across a broad swath of central Blantyre, with
dozens of wards, specialised facilities, and administrative buildings,
linked together by a bewildering maze of covered walkways. A hive of
around-the-clock activity, Queen's bustles from dawn to dusk with a
constant stream of patients, drawn from across the country, and
maintains a city-like feel even after dark as staff and stay-in family
members cook and care for their in-patient dependants. Yet, for the
first-time visitor, it's not the architecture or the crowds that are the
most striking when entering the grounds, but the smell. Immediately
noticeable, even from the road outside the hospital, Queen's is
smothered by an acrid-smelling, white smoke, which hangs over the
grounds day and night: the product of several constantly burning fires
spread across the campus. Breathing this air, when a significant amount
of burning is occurring, primarily at night when the smoke is less
visible, can be incredibly uncomfortable. Moreover, the tropical
architecture, designed to let air and light circulate, means that even
indoors, within patient wards, surgery, and treatment areas, the air
quality can also be unbearable. Poor air quality associated with
burning, through released particulate matter (PM), has been linked to
multiple negative health outcomes like asthma, heart disease, heart
failure, stroke, and cancer among others (Anderson et al., 2012; Bell &
HEI Health Review Committee, 2012). In a space of healing and recovery,
why is such burning occurring, and how can it continue to persist?

# Material and methods

# Theory/calculation (not applicable)

# Results and Discussion

# Conclusions

# Appendix

If there is more than one appendix, they should be identified as A, B,
etc. Formulae and equations in appendices should be given separate
numbering: Eq. (A.1), Eq. (A.2), etc.; in a subsequent appendix, Eq.
(B.1) and so on. Similarly for tables and figures: Table A.1; Fig. A.1,
etc.

# Template instructions

Please make sure that your manuscript follows the guidelines in the
Guide for Authors of the relevant journal. It is not necessary to
typeset your manuscript in exactly the same way as an article, unless
you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see
<https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

Here are two sample references: @Feynman1963118 @Dirac1953888.

By default, natbib will be used with the `authoryear` style, set in
`classoption` variable in YAML. You can sets extra options with
`natbiboptions` variable in YAML header. Example

```         
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at
<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>.
To use one of these, add it in the header using, for example,
`biblio-style: model1-num-names`.

## Using CSL

If `cite-method` is set to `citeproc` in `elsevier_article()`, then
pandoc is used for citations instead of `natbib`. In this case, the
`csl` option is used to format the references. By default, this template
will provide an appropriate style, but alternative `csl` files are
available from <https://www.zotero.org/styles?q=elsevier>. These can be
downloaded and stored locally, or the url can be used as in the example
header.

# Equations

Here is an equation: $$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple
example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {.unnumbered}