### 2023-04-05 manuscript/manuscript.qmd
Transfer introduction, update references

---
title: Short Paper
subtitle: A Short Subtitle
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliations: 
        - id: some-tech
          name: Some Institute of Technology
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Bob Security
    email: bob@example.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    note: |
      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  - name: Cat Memes
    email: cat@example.com
    affiliations:
        - ref: another-u
    note: Yet another author footnote.
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: [bibliography.bib, introduction.bib, man-qech-air-quality.bib]
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Habitat International
      formatting: review
      model: 3p
      cite-style: authoryear
      highlights:
        - Highlight 1
        - Highlight 2
        - Highlight 3
editor: 
  markdown: 
    wrap: 72
---

# Introduction

Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
(QECH), or Queen's, as it's known locally, is the country's largest
public hospital. Designed in the Tropical Modernism style of
architecture popular within Africa in the late colonial period, Queen's
open-plan form sprawls across a broad swath of central Blantyre, with
dozens of wards, specialised facilities, and administrative buildings,
linked together by a bewildering maze of covered walkways. A hive of
around-the-clock activity, Queen's bustles from dawn to dusk with a
constant stream of patients, drawn from across the country, and
maintains a city-like feel even after dark as staff and stay-in family
members cook and care for their in-patient dependants. Yet, for the
first-time visitor, it's not the architecture or the crowds that are the
most striking when entering the grounds, but the smell. Immediately
noticeable, even from the road outside the hospital, Queen's is
smothered by an acrid-smelling, white smoke, which hangs over the
grounds day and night: the product of several constantly burning fires
spread across the campus. Breathing this air, when a significant amount
of burning is occurring, primarily at night when the smoke is less
visible, can be incredibly uncomfortable. Moreover, the tropical
architecture, designed to let air and light circulate, means that even
indoors, within patient wards, surgery, and treatment areas, the air
quality can also be unbearable. Poor air quality associated with
burning, through released particulate matter (PM), has been linked to
multiple negative health outcomes like asthma, heart disease, heart
failure, stroke, and cancer among others [@anderson2012clearing; @bell2012assessment].
In a space of healing and recovery,
why is such burning occurring, and how can it continue to persist?

Queen’s, simply, produces an immense amount of waste that it is unable to efficiently manage. The dozens of wards, offices, and kitchens, serving the hundreds of patients and staff, generates considerable infectious medical waste, mixed domestic and office waste, and kitchen waste, which needs to be removed and disposed of daily. Some individual wards implement waste separation, binning medical wastes separately from general waste, however separation ends at the point of collection. Queen’s is not served by regular municipal waste collection. What municipal collection that is done is ad-hoc, sporadic, and based on the hospital’s available financial resources. Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and burnt at the hospital’s incinerator, located at a central point within the hospital campus  [@kalina2020this]. However, the vast majority of the deposited waste is not incinerated, as a new incinerator provided by the Ministry of Health in late 2019, is only able to handle a small percentage of the hospital’s waste. Furthermore, the incinerator is frequently inoperable, under repair, or without power,. Rather, the waste is openly burned, in a constantly smouldering pile to the side of the incinerator building, sending up the aforementioned clouds of smoke, which choke the grounds (Figure 1). Innumerable other, smaller, fires, scattered across the QECH campus (Figure 2) contribute their part, as grounds and maintenance staff habitually burn leaves and other garden refuse. Furthermore, caregivers who reside separately on hospital grounds, cook for themselves throughout the day and night using biomass (charcoal or wood), and burn their own domestic waste. As a result, air quality at the hospital is a constant source of discussion, with patients, staff, and caregivers struggling to cope, when just breathing is, as the quote used in the title of this article, ‘unbearable.’

Challenges with solid waste management (SWM) are not unique to Queen’s, and remain persistent globally, particularly within African nations, and the Global South more broadly  [@bundhoo2018solid; @sergekubanza2022exploring; @muheirwe2022paradox; @srivastava2015urban]. Explanations for these disparities have centred on the on waste generation outstripping waste management infrastructure [@muheirwe2022paradox], and the high costs of waste management systems, which often prove unaffordable for many low-and-middle-income countries (LMICs) [@guerrero2013solid; @sergekubanza2022exploring]. Waste collection is an important step within SWM systems, and is a common barrier within LMIC contexts [@bundhoo2018solid]. Waste inequalities between high-income countries (HICs) and low-income countries (LICs) is often most pernicious and visible at the point of collection, with some HICs achieving 100% collection rates, while some African nations have rates well below 50% [@shi2021unbalanced]. Inequalities persist beyond collection, however. Even once waste is collected, municipalities in LMICs may burn or dump it in non-sanitary landfills due to the absence of further treatment or disposal options [@cook2021globala]. According to @kaza2018what, 90% of waste in LIC is disposed of in unregulated dumps or openly burned, while the quantity of waste generation in these countries is expected to triple by 2050. As a result, open waste burning, in nations like Malawi, can be expected to become more prevalent.

```{r}
# Figure 1: Open burning at QECH, with the old incinerator building in the background (Authors).
knitr::include_graphics(here::here("manuscript/img/image-replace-tree.png"))
```

Although hospital waste is among the common types of solid waste openly burned in LICs (in addition to municipal solid waste, sewage sludge, market or commercial waste, agricultural residues), the open burning of hospital waste has not garnered as much discussion as the burning of other waste fractions [@cook2021globala]. The World Health Organisation (WHO) defines hospital, or health-care waste, as waste generated by health-care activities, of which approximately 85% is general, non-hazardous waste, but the remaining 15% can be considered hazardous, and may be toxic, infectious, or radioactive. These infectious fractions are commonly incinerated in both high and middle-income countries, however in the absence of capacity, or costly incineration or autoclave infrastructure, open burning is often a last resort option for disposal in LIC contexts [@who2018healthcare]. The open burning of biomedical waste may reduce infection risk from potentially harmful pathogens, however, such waste may contain sharps, radioactive waste, mercury-containing instruments, and plastics, the open burning of which may result in the emission of dioxins, furans, and toxic particulate matter (PM) [@who2018healthcare; @cook2021globala].

Poor air quality in in African cities can be attributed to multiple factors, including vehicular emissions, dust, or the use of low-quality fuels for cooking, lighting, and heating, however, the open burning of waste has been identified as one of the key contributors to air pollution within urban areas [@amegah2018proliferation; @bulto2020impact].  Although open trashing burning is a problem in LICs broadly, for instance, @wiedinmyer2014global estimate that open trash burning contributes to 29% of the total global PM2.5 anthropogenic emissions, it is of particular concern in Sub-Saharan African (SSA) cities, which are home to 19 of the world’s 50 biggest dumpsites  (United Nations Development Programme, 2021). Particulate matter, or PM, is not specific to trash burning, but rather the product of any type of incomplete combustion and is one of the most commonly measured indicators for air quality. PM is measured in terms of its particle size and generally classified into PM with particle sizes less than 10μm (PM10) and those with sizes less than 2.5μm (PM2.5), while one way of measuring the severity of air pollution is by expressing the mass of PM per volume of air (μg PMx/m3 air). PM has been linked to multiple negative health outcomes. Amongst particulate air pollution, PM10 and PM2.5 are of particular concern as they can penetrate deep into the lungs, and their exposure has been associated with asthma, heart disease, heart failure, stroke, and cancer [@mohapatra2014effect; @xing2016impact]. As a result, the health impacts of ambient air particle pollution may be significant, especially for the most vulnerable. For instance, the @ihme2019ambient estimates PM related air pollution resulted in 4.14 million deaths worldwide (in 2019), of which, 394,000 occurred in Africa. Yet, despite its importance, the air quality and health picture in Africa is worryingly incomplete. For instance, the WHO only collects data from 10 African countries (covering 39 cities) [@worldhealthorganization2016ambient]. Likewise, robust, epidemiological data from and for the African context is also lacking, though emerging data suggests strong correlations between PM and poor health outcomes [@coker2018narrative; @heft-neal2018robust]. Moving forward, it is not just the air at Queen’s which lacks clarity.

Utilising a mixed-methods approach, including a network of custom-designed air quality monitors and qualitative fieldwork with hospital staff, patients, and caregivers, the purpose of this investigation was to gain a multi-dimensional understanding of the impacts of the open burning of the various solid and organic waste fractions at QECH. Specifically, this work aimed to intensively and longitudinally measure the air quality, in particulate matter (PM), at multiple locations at and surrounding the central burning point. Furthermore, we also aimed to qualitatively understand how affected individuals perceive air quality at QECH and understand potential health impacts.

This research responds to a several specific gaps within the body of academic air quality literature: Although there has been ample discussion of air quality challenges in African cities [@mbow-diokhane2019air; @miralvarez2020scoping; @petkova2013particulate; @singh2021aira], there is a paucity of data on air quality impacts linked to trash burning. In particular, within African cities, there has been a total dearth of scholarship on the burning of medical waste, or on air quality within hospital contexts. Furthermore, while the voices of the citizens of Northern cities [@deguen2012new; @howel2002urban; @reames2019people] have been well documented on localised air quality issues, the voices of African urban dwellers have not been given the same consideration, and there remains a pressing need to understand how these populations experience the impacts of the open burning of trash within their communities and understand potential risks.

Findings suggest that particulate matter concentrations are routinely above the WHO guidelines and are especially worrisome at several locations. Over the course of the two-month study, the hazardous limits for both PM10 and PM2.5 were exceeded at all locations except for the Lions Sight hospital. The limits were exceeded fewer than 50 times at five of the locations, but the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the hazardous limits.  The extremely hazardous air quality at the Guardian shelter is mostly from cooking over open fires, while the PM at the Lighthouse Clinic is a function of being located directly adjacent to the smouldering waste pile.

Interviews show that patients, staff, and caregivers alike, are keenly aware that the air quality at Queen’s is poor, with most respondents reporting frequent respiratory-related illness. Many also linked the smoke to potential long-term health complications and expressed a belief that the pollution could be contributing to other, more potentially life-threatening diseases, such as asthma, cancer, and tuberculosis. Moreover, the tropical design of hospital buildings have rendered most coping mechanisms ineffective, with most staff only finding relief at home: relief that is not available to the hundreds of patients and caregivers who sleep at QECH or are unable to leave.

In many contexts, waste burning is a necessary, but sporadic event with limited exposure.  However, in the context of a large hospital, serving the sickest and poorest, this study shows that the consistent, and highly toxic smoke produced is an unrelenting and unnecessary burden that must be addressed, lest the patients leave in worse condition than in which they arrived.

# Material and methods

# Theory/calculation (not applicable)

# Results and Discussion

# Conclusions

# Appendix

If there is more than one appendix, they should be identified as A, B,
etc. Formulae and equations in appendices should be given separate
numbering: Eq. (A.1), Eq. (A.2), etc.; in a subsequent appendix, Eq.
(B.1) and so on. Similarly for tables and figures: Table A.1; Fig. A.1,
etc.

# Template instructions

Please make sure that your manuscript follows the guidelines in the
Guide for Authors of the relevant journal. It is not necessary to
typeset your manuscript in exactly the same way as an article, unless
you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see
<https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

Here are two sample references: @Feynman1963118 @Dirac1953888.

By default, natbib will be used with the `authoryear` style, set in
`classoption` variable in YAML. You can sets extra options with
`natbiboptions` variable in YAML header. Example

```         
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at
<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>.
To use one of these, add it in the header using, for example,
`biblio-style: model1-num-names`.

## Using CSL

If `cite-method` is set to `citeproc` in `elsevier_article()`, then
pandoc is used for citations instead of `natbib`. In this case, the
`csl` option is used to format the references. By default, this template
will provide an appropriate style, but alternative `csl` files are
available from <https://www.zotero.org/styles?q=elsevier>. These can be
downloaded and stored locally, or the url can be used as in the example
header.

# Equations

Here is an equation: $$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple
example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {.unnumbered}