### 2023-04-06 manuscript/manuscript.qmd
Fix units

---
title: "'It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi"
author:
  - name: Elizabeth Tilley
    orcid: 0000-0002-2095-9724
    email: tilleye@ethz.ch
    affiliations: 
        - id: ethz
          name: ETH Zurich
          department: Department of Mechanical and Process Engineering
          address: Global Health Engineering, Clausiusstrasse 37
          city: Zurich
          country: Switzerland
          postal-code: 8092
    attributes:
        corresponding: true
  - name: Hope Chilunga
    orcid: 0000-0003-4216-6319
    email: hchilunga@poly.ac.mw
    affiliations: 
        - id: mubas
          name: Malawi University of Business and Applied Sciences
          department: Department of Environmental Health
          city: Blantyre
          country: Malawi
  - name: Jonathan Kwangulero
    orcid: 0000-0002-4050-5121
    email: kwangerlo@gmail.comm
    affilations:
        - ref: mubas
  - name: Lars Schöbitz
    orcid: 0000-0003-2196-5015
    email: lschoebitz@ethz.ch
    affiliations:
        - ref: ethz
  - name: Heiko Heilgendorff
    email: me@me.com
    affiliations: 
        - id: ukzn
          name: University of KwaZulu-Natal
          department: School of Mathematics, Statistics and Computer Science
          city: Durban
          country: South Africa
  - name: Marc Kalina
    orcid: 0000-0002-6335-3845
    email: mkalina@ethz.ch
    affiliations:
        - ref: ethz
abstract: |
  Blantyre, Malawi’s Queen Elizabeth Central Hospital (QECH), or Queen’s, as it’s known locally, is the country’s largest public hospital. However, Queen’s is not served by regular municipal waste collection. What municipal collection that is done is ad-hoc, sporadic, and based on the hospital’s available financial resources. Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and openly burned, in several constantly smouldering piles, sending up clouds of smoke.  Speaking directly to an identified knowledge gap on air quality impacts linked to trash burning and the paucity of African urban dwellers’ voices on air quality issues, this study employed a mixed-methods approach to both quantitatively measure the air quality around QECH, and to qualitatively investigate the perceived impacts amongst staff and caregivers. Low-cost sensors measuring particulate matter (PM) with particle sizes less than 10 µm (PM~10~) and less than 2.5 µm (PM~2.5~), expressed as the mass of PM per volume of air (µg PMx/m^3^ air) were recorded every 5 minutes at 8 locations across the QECH for 2 months. Qualitative data collection consisted of 57 interviews with patients, caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators).  Our results show that safe air quality thresholds are consistently exceeded across space and time and that the most problematic air quality surrounds the shelter for caregivers and those receiving treatment for HIV/AIDs. Moreover, staff and visitors are severely impacted by the poor air quality within the space, but  feel powerless to make changes or address complaints. Waste management interventions are desperately needed lest the patients who arrive at Queen’s leave with more health issues than the ones with which they arrived. 
keywords: 
  - waste management
  - healthcare waste
  - trash burning
  - air quality
  - Malawi
  - urbanisation
date: last-modified
bibliography: man-qechairquality.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Habitat International
      formatting: review
      model: 3p
      cite-style: authoryear
      highlights:
        - Low-cost monitors were used to assess air quality at Malawi’s largest hospital
        - Open waste burning is physically extensive and temporally frequent across the hospital campus
        - Air quality measured as PM~2.5~ and PM~10~ consistently exceeded even the most hazardous levels 
        - The most exposed locations were where guardians of patients stay and the HIV clinic
execute: 
  echo: false
  warning: false
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: console
---

```{r setup}

# Source data preparation script
source(here::here("R/01-data-preparation.R"))

```

# Introduction

Built in 1964, Malawi's year of independence, Queen Elizabeth Central Hospital
(QECH), or Queen's, as it's known locally, is the country's largest
public hospital. Designed in the Tropical Modernism style of
architecture popular within Africa in the late colonial period, Queen's
open-plan form sprawls across a broad swath of central Blantyre, with
dozens of wards, specialised facilities, and administrative buildings,
linked together by a bewildering maze of covered walkways. A hive of
around-the-clock activity, Queen's bustles from dawn to dusk with a
constant stream of patients, drawn from across the country, and
maintains a city-like feel even after dark as staff and stay-in family
members cook and care for their in-patient dependants. Yet, for the
first-time visitor, it's not the architecture or the crowds that are the
most striking when entering the grounds, but the smell. Immediately
noticeable, even from the road outside the hospital, Queen's is
smothered by an acrid-smelling, white smoke, which hangs over the
grounds day and night: the product of several constantly burning fires
spread across the campus. Breathing this air, when a significant amount
of burning is occurring, primarily at night when the smoke is less
visible, can be incredibly uncomfortable. Moreover, the tropical
architecture, designed to let air and light circulate, means that even
indoors, within patient wards, surgery, and treatment areas, the air
quality can also be unbearable. Poor air quality associated with
burning, through released particulate matter (PM), has been linked to
multiple negative health outcomes like asthma, heart disease, heart
failure, stroke, and cancer among others [@anderson2012clearing;
@bell2012assessment]. In a space of healing and recovery, why is such
burning occurring, and how can it continue to persist?

Queen's, simply, produces an immense amount of waste that it is unable
to efficiently manage. The dozens of wards, offices, and kitchens,
serving the hundreds of patients and staff, generates considerable
infectious medical waste, mixed domestic and office waste, and kitchen
waste, which needs to be removed and disposed of daily. Some individual
wards implement waste separation, binning medical wastes separately from
general waste, however separation ends at the point of collection.
Queen's is not served by regular municipal waste collection. What
municipal collection that is done is ad-hoc, sporadic, and based on the
hospital's available financial resources. Rather, most hospital waste
(infectious and non-infectious) is gathered by grounds staff and burnt
at the hospital's incinerator, located at a central point within the
hospital campus[^1] [@kalina2020this]. However, the vast majority of the
deposited waste is not incinerated, as a new incinerator provided by the
Ministry of Health in late 2019, is only able to handle a small
percentage of the hospital's waste. Furthermore, the incinerator is
frequently inoperable, under repair, or without power[^2]. Rather, the
waste is openly burned, in a constantly smouldering pile to the side of
the incinerator building, sending up the aforementioned clouds of smoke,
which choke the grounds (@fig-open-burning). Innumerable other, smaller, fires,
scattered across the QECH campus (@fig-map-sensors) contribute their part, as
grounds and maintenance staff habitually burn leaves and other garden
refuse. Furthermore, caregivers[^3] who reside separately on hospital
grounds, cook for themselves throughout the day and night using biomass
(charcoal or wood), and burn their own domestic waste. As a result, air
quality at the hospital is a constant source of discussion, with
patients, staff, and caregivers struggling to cope, when just breathing
is, as the quote used in the title of this article, 'unbearable.'

[^1]: Sharps are handled and disposed of separately.

[^2]: There was an older incinerator on the grounds which failed in
2017. Between 2017 and 2021, when the new incinerator was
commissioned, nothing was incinerated at all.

[^3]: Patients are required to bring a caregiver for in-patient
hospitals stays, usually a family member, who assist with feeding,
bathing, and all other non-medical services.

Challenges with solid waste management (SWM) are not unique to Queen's,
and remain persistent globally, particularly within African nations, and
the Global South more broadly [@bundhoo2018solid;
@sergekubanza2022exploring; @muheirwe2022paradox; @srivastava2015urban].
Explanations for these disparities have centred on the on waste
generation outstripping waste management infrastructure
[@muheirwe2022paradox], and the high costs of waste management systems,
which often prove unaffordable for many low-and-middle-income countries
(LMICs) [@guerrero2013solid; @sergekubanza2022exploring]. Waste
collection is an important step within SWM systems, and is a common
barrier within LMIC contexts [@bundhoo2018solid]. Waste inequalities
between high-income countries (HICs) and low-income countries (LICs) is
often most pernicious and visible at the point of collection, with some
HICs achieving 100% collection rates, while some African nations have
rates well below 50% [@shi2021unbalanced]. Inequalities persist beyond
collection, however. Even once waste is collected, municipalities in
LMICs may burn or dump it in non-sanitary landfills due to the absence
of further treatment or disposal options [@cook2021globala]. According
to @kaza2018what, 90% of waste in LIC is disposed of in unregulated
dumps or openly burned, while the quantity of waste generation in these
countries is expected to triple by 2050. As a result, open waste
burning, in nations like Malawi, can be expected to become more
prevalent.

```{r}
#| label: fig-open-burning
#| fig-cap: Open burning at QECH, with the old incinerator building in the background (Authors).
#| fig-align: center
#| out-width: 80%

knitr::include_graphics(here::here("manuscript/img/open-waste-burning-incinerator-qech.jpg"))
```

Although hospital waste is among the common types of solid waste openly
burned in LICs (in addition to municipal solid waste, sewage sludge,
market or commercial waste, agricultural residues), the open burning of
hospital waste has not garnered as much discussion as the burning of
other waste fractions [@cook2021globala]. The World Health Organisation
(WHO) defines hospital, or health-care waste, as waste generated by
health-care activities, of which approximately 85% is general,
non-hazardous waste, but the remaining 15% can be considered hazardous,
and may be toxic, infectious, or radioactive. These infectious fractions
are commonly incinerated in both high and middle-income countries,
however in the absence of capacity, or costly incineration or autoclave
infrastructure, open burning is often a last resort option for disposal
in LIC contexts [@who2018healthcare]. The open burning of biomedical
waste may reduce infection risk from potentially harmful pathogens,
however, such waste may contain sharps, radioactive waste,
mercury-containing instruments, and plastics, the open burning of which
may result in the emission of dioxins, furans, and toxic particulate
matter (PM) [@who2018healthcare; @cook2021globala].

Poor air quality in in African cities can be attributed to multiple
factors, including vehicular emissions, dust, or the use of low-quality
fuels for cooking, lighting, and heating, however, the open burning of
waste has been identified as one of the key contributors to air
pollution within urban areas [@amegah2018proliferation;
@bulto2020impact]. Although open trashing burning is a problem in LICs
broadly, for instance, @wiedinmyer2014global estimate that open trash
burning contributes to 29% of the total global PM~2.5~ anthropogenic
emissions, it is of particular concern in Sub-Saharan African (SSA)
cities, which are home to 19 of the world's 50 biggest dumpsites (United
Nations Development Programme, 2021). Particulate matter, or PM, is not
specific to trash burning, but rather the product of any type of
incomplete combustion and is one of the most commonly measured
indicators for air quality. PM is measured in terms of its particle size
and generally classified into PM with particle sizes less than 10 µm
(PM~10~) and those with sizes less than 2.5 µm (PM~2.5~), while one way of
measuring the severity of air pollution is by expressing the mass of PM
per volume of air (µg PMx/m^3^ air). PM has been linked to multiple
negative health outcomes. Amongst particulate air pollution, PM~10~ and
PM~2.5~ are of particular concern as they can penetrate deep into the
lungs, and their exposure has been associated with asthma, heart
disease, heart failure, stroke, and cancer [@mohapatra2014effect;
@xing2016impact]. As a result, the health impacts of ambient air
particle pollution may be significant, especially for the most
vulnerable. For instance, the @ihme2019ambient estimates PM related air
pollution resulted in 4.14 million deaths worldwide (in 2019), of which,
394,000 occurred in Africa. Yet, despite its importance, the air quality
and health picture in Africa is worryingly incomplete. For instance, the
WHO only collects data from 10 African countries (covering 39 cities)
[@worldhealthorganization2016ambient]. Likewise, robust, epidemiological
data from and for the African context is also lacking, though emerging
data suggests strong correlations between PM and poor health outcomes
[@coker2018narrative; @heft-neal2018robust]. Moving forward, it is not
just the air at Queen's which lacks clarity.

Utilising a mixed-methods approach, including a network of
custom-designed air quality monitors and qualitative fieldwork with
hospital staff, patients, and caregivers, the purpose of this
investigation was to gain a multi-dimensional understanding of the
impacts of the open burning of the various solid and organic waste
fractions at QECH. Specifically, this work aimed to intensively and
longitudinally measure the air quality, in particulate matter (PM), at
multiple locations at and surrounding the central burning point.
Furthermore, we also aimed to qualitatively understand how affected
individuals perceive air quality at QECH and understand potential health
impacts.

This research responds to a several specific gaps within the body of
academic air quality literature: Although there has been ample
discussion of air quality challenges in African cities
[@mbow-diokhane2019air; @miralvarez2020scoping; @petkova2013particulate;
@singh2021aira], there is a paucity of data on air quality impacts
linked to trash burning. In particular, within African cities, there has
been a total dearth of scholarship on the burning of medical waste, or
on air quality within hospital contexts. Furthermore, while the voices
of the citizens of Northern cities [@deguen2012new; @howel2002urban;
@reames2019people] have been well documented on localised air quality
issues, the voices of African urban dwellers have not been given the
same consideration, and there remains a pressing need to understand how
these populations experience the impacts of the open burning of trash
within their communities and understand potential risks.

Findings suggest that particulate matter concentrations are routinely
above the WHO guidelines and are especially worrisome at several
locations. Over the course of the two-month study, the hazardous limits
for both PM~10~ and PM~2.5~ were exceeded at all locations except for the
Lions Sight hospital. The limits were exceeded fewer than 50 times at
five of the locations, but the monitors at the Lighthouse Clinic and the
Guardian Shelter recorded more than 50 instances above the hazardous
limits. The extremely hazardous air quality at the Guardian shelter is
mostly from cooking over open fires, while the PM at the Lighthouse
Clinic is a function of being located directly adjacent to the
smouldering waste pile.

Interviews show that patients, staff, and caregivers alike, are keenly
aware that the air quality at Queen's is poor, with most respondents
reporting frequent respiratory-related illness. Many also linked the
smoke to potential long-term health complications and expressed a belief
that the pollution could be contributing to other, more potentially
life-threatening diseases, such as asthma, cancer, and tuberculosis.
Moreover, the tropical design of hospital buildings have rendered most
coping mechanisms ineffective, with most staff only finding relief at
home: relief that is not available to the hundreds of patients and
caregivers who sleep at QECH or are unable to leave.

In many contexts, waste burning is a necessary, but sporadic event with
limited exposure. However, in the context of a large hospital, serving
the sickest and poorest, this study shows that the consistent, and
highly toxic smoke produced is an unrelenting and unnecessary burden
that must be addressed, lest the patients leave in worse condition than
in which they arrived.

# Material and methods

This study utilised a mixed-methods approach to both quantitatively
measure the air quality at eight locations around QECH, and to
qualitatively investigate the perceived impacts amongst staff and
caregivers. Although the work was conducted over one sustained period in
late 2019, it must be contextualised within five years' experience of
research and activism within QECH by the authors. All relevant
permissions were obtained from QECH beforehand, through consultations
with administration and staff, and the research was approved by the
National Committee On Research in The Social Sciences And Humanities in
Malawi, Protocol NO. P.03/19/356.

## Study site

The sensors were located within the QECH campus in consultation with
hospital management. The goal was to locate sensors across the campus
with a range of distances and directions from the incinerator, though
the final decision was based on accessibility, the permission of the
unit's head, and convenience. Each of the locations is briefly described
below and shown in Figure 2:

-   **Lions Sight First Eye Hospital**, often shortened to **Lions
Sight**, is the largest eye hospital in Malawi and serves as the
main teaching eye hospital for the Kamuzu University of Health
Sciences (formerly College of Medicine). It is staffed by 5
consultants and a team of clinical officers and nurses. It provides
a mix of public (no cost) and private (at cost) services.

-   **The Blantyre Malaria Project (BMP)**, established by Professors
Terrie Taylor and Malcolm Molyneux, has carried out clinical
research and patient care in the area of paediatric malaria,
specifically cerebral malaria, for more than 30 years[^4]. Its
research infrastructure includes an administrative team, an
inpatient research ward, an MRI centre, and an outpatient research
clinic in Ndirande township and in other districts outside of
Blantyre.

-   **Ward 6B** is a male ward for trauma and orthopaedic patients. It
was established as part of the original hospital design. It is a
60-bed ward with a dedicated nursing station and treatment rooms.

-   **The Blantyre Lighthouse Trust Clinic**, also known as **Umodzi
Family Centre** or simply, the **Lighthouse Clinic**, is one of four
operating across Malawi. The clinics work in collaboration with the
Ministry of Health (MOH) to provide integrated HIV testing,
treatment and care for people living with HIV.

-   The **Moyo Nutritional Rehabilitation and Research Unit**, known on
campus as **uMoyo**, is a 57-bed nutritional rehabilitation unit
(NRU) for treating infants and children with severe malnutrition and
acute illnesses. It is one of 104 operational NRUs in Malawi. It
also has an outpatient therapeutic feeding program (OTP) for
children with malnutrition who can be treated outside of the
hospital.

-   **Mercy James Institute for Pediatric Surgery and Intensive Care
(MJC)**, shortened on campus to **Mercy James**, was opened in 2017
by the NGO Raising Malawi, which was founded by Madonna. With 3
operating rooms and 50 beds, it is the first and only first
dedicated paediatric surgery hospital in Malawi. Prior to its
opening, QECH had fewer than 10 intensive care beds.

-   The **Administration Building** is the operational hub of the
campus, containing mostly offices and meeting rooms.

-   When a patient arrives at QECH, she or he will usually arrive with a
guardian: someone to cook for them, buy medicine, do their laundry,
and help them bathe. The **Guardian Shelter** is a gender-separated
concrete floor shelter for sleeping along with a cooking pavilion
and toilet/shower block are maintained by a the local NGO Chira
fund. Though simple, the facility is secure and offers one of the
only free/accessible toilet facilities for visitors on the QECH
campus.

[^4]: BMP is credited with the development of the Blantyre Coma Score,
which is used widely.

```{r}
#| label: prepare-sensor-map

# Icon from: https://github.com/apancik/public-domain-icons/blob/master/dist/symbol%20fire.svg

fire_icon <- makeIcon(iconUrl = here::here("manuscript/img/symbol-fire.png"),
                      iconWidth = 50,
                      iconHeight = 50)

# Create a palette that maps factor levels to colors
pal <- colorFactor(c("#EE4B2B", "navy"), domain = c("sensor", "fire"))

fire_loc <- locations  |>
  filter(type == "fire")

sensor_loc <- locations  |>
  filter(type == "sensor")

qech_sensor_map <- leaflet(sensor_loc)  |>
  setView(lng = 35.021071999805755,
          lat = -15.803269709021687,
          zoom = 18)  |>
  #addProviderTiles(providers$Stamen)  |>
  addTiles()  |>
  addMarkers(
    lng = fire_loc$long,
    lat = fire_loc$lat,
    label = fire_loc$location,
    icon = ~fire_icon,
    labelOptions = labelOptions(noHide = T,
                                textsize = "15px",
                                direction = "bottom",
                                offset = c(0, 20),
                                style = list(
                                  "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
                                  "border-color" = "rgba(0,0,0,0.5)"
                                  #"font-style" = "italic"
                                ))
  )  |>
  addCircleMarkers(
    lng = ~long,
    lat = ~lat,
    label = ~location,
    radius = 16,
    stroke = FALSE,
    fillOpacity = 1,
    color = ~pal(type),
    #icon = ~fire_icon,
    labelOptions = labelOptions(noHide = T,
                                textsize = "15px",
                                direction = "bottom",
                                offset = c(0, 10),
                                style = list(
                                  "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
                                  "border-color" = "rgba(0,0,0,0.5)"
                                  #"font-style" = "italic"
                                )))

qech_sensor_map |> 
  mapshot(file = here::here("manuscript/figs/map-sensors-qech-blantyre.png"))
```

```{r}
#| label: fig-map-sensors
#| fig-cap: Sensor locations within QECH
#| fig-align: center
#| out-width: 80%

# Figure 2: Sensor locations within QECH
knitr::include_graphics(here::here("manuscript/figs/map-sensores-qech-blantyre.png"))
```

## Waste management

As the largest public hospital in the country, Queen's produces an
immense amount of waste. Hospital staff, such as doctors and nurses,
primarily generate and handle medical waste (dry, wet, sharps), which is
disposed of in separate waste bins labelled accordingly within the ward.
Hospital visitors (caretakers and patients) generate food waste and are
advised by staff to discard such waste in waste bins outside, on
hospital grounds. Cleaning staff are responsible for emptying bins and
maintaining the cleanliness of indoor and outdoor spaces. Collected bags
are temporarily stored in the sluice room, and cleaners during the night
shift are responsible for transporting the waste on wheelchairs or
trolleys to the incinerator for burning by waste disposers.

Yet, according to respondents and our own observations, waste management
procedures within the hospital are fraught with challenges which impact
the implementation of best practices. These include the lack of a
coherent hospital-wide waste management policy, insufficient waste
management training for staff, and a persistent lack of material and
financial resources. For instance, although hospital administration was
able to articulate a set of standard operating principles (SoPs) for
waste management, below the administration, doctors and nurses were
generally limited in their understanding about its contents.
Furthermore, cleaners, who are responsible for day-to-day cleansing and
waste disposal, yet rate at the bottom of the institutional hierarchy at
Queens, were unable to access or analyse hospital policy, and generally
tended to rely on "experience" and "human judgement" when handling
potentially hazardous waste. Moreover, although nearly all doctors and
nurses had prior education on waste management best practices, cleaners
began their waste handling duties without any prior waste management
experience, and although there is an orientation programme for new
hires, there is no regular refresher training or capacity building
regarding waste for existing staff, at any level.

Despite its national prominence, and international reputation for
academic medical research, Queen's is plagued by constant shortages of
human and material resources. Yet, these shortages are not shared
equally across the space, with a sharp division in cleanliness,
resources, and quality of care between public wards, supported by the
Malawi Ministry of Health, and the private wads with international
funding, such as the aforementioned Mercy James Institute. For staff,
these shortages have manifested in poor salaries, missed pay cheques,
failing infrastructure and limited supplies [@kalina2020this]. Regarding
waste management, these shortages have manifested in a chronic shortage
of black bags, waste containers and cleaning staff. As a result, due to
a lack of bags and storage containers, there is minimal segregation of
waste, and highly infectious items (bandages, gloves, syringes, and
other sharps) are disposed of along with non-hazardous waste (food,
newspapers, packaging). As a result, rather than facilitating different
disposal pathways for hazardous and non-hazardous waste, all waste is
mixed, necessitating more burning than would be necessary if
non-hazardous waste was diverted, and contributing to workplace risk,
with reports of accidental exposure to potentially infected waste,
including accidental pokes from improperly disposed of sharps, common
amongst hospital staff. Regardless, separation ends at the point of
collection. The hospital is not serviced by regular municipal waste
collection, and as a result, nearly all of the waste produced by the
hospital, hazardous and non-hazardous, has to be disposed of on-site,
either within the limited capacity of the hospital incinerator, or more
commonly, through open burning. As a result, the constantly smouldering
pile of waste puts forth a continuous cloud of grey smoke, which
mingling with the dozens of other fires on the grounds, from cooking
fires and burning garden waste, blankets Queen's in a permanent cloud of
foul smelling haze.

## Measuring Air Quality

### Hardware and Software

The PM monitoring device centres around a Raspberry Pi 3 Model B
single-board computer complete with an operating system and storage
space. The added advantage of storage space is that the data are not
lost if the internet connection breaks down. Because the Pi can only
read digital signals, we needed to include an analog-to-digital (ADC)
converter. The amount of dust is measured by a Nova PM SDS011 High
Precision Laser sensor. This sensor measures particles at 2.5 and 10
µm in diameter in mg/m^3^. A list of the specific hardware and
software components can be found in Appendix B.

### Installation

The choice of where the sensors were installed was done in collaboration
with the QECH administration, but the position of the sensors at the
individual buildings were decided by the research team. Four air quality
sensors were installed on the outside of buildings (Guardian Shelter,
Mercy James, Malaria Project, Lighthouse Clinic) while the remaining
four sensors were installed inside (Ward 6B, Administration, Lions
Sight, uMoyo). All of the sensors were mounted to a wall at a height of
approximately two metres from the ground with the help of QECH
maintenance department personnel. This height was chosen to prevent the
public from tampering with the equipment when they visit the hospital.
Additionally, we assured that each pipe that inhaled ambient air was
freely protruding in the building (room) in order to capture the air
quality.

Because connecting each unit to a power supply was not possible, each
unit was equipped with an external battery (5000mAh) which powered the
sensor unit for three days continuously. To avoid any down time, we
changed the batteries every two days. Data were also collected from the
sensor at the time of battery replacement. The sensors were capable of
being connected to WIFI, a feature which enabled us to download the data
into our computers wirelessly.

### Qualitative Methods

Qualitative data collection consisted of 26 interviews with caregivers
and hospital staff (including janitorial and maintenance staff, nurses,
doctors, and administrators) conducted in October and November 2019, as
well as an additional 31 interviews conducted with staff, patients, and
caregivers in January and February 2021. Interviews were
semi-structured, and included questions on hospital waste management
practices, perceived best practices, and workplace health, safety, and
risk. However, the semi-structured nature of the interviews allowed for
the participants to raise topics of interest, and for unexpected themes
to emerge. These interviews are supplemented by participatory
observation recorded within detailed field notes, and supported by
nearly a decade of sustained involvement and activism by the authors in
the waste management practices of the hospital.

Interview respondents were chosen using a purposive or judgement
sampling regimen, i.e. a subjective sampling method in which respondents
are selected based on their ability to effectively contribute to the
study's research objectives [@kitchin2013conducting]. All respondents
were purposively selected based on availability, willingness to provide
written, informed consent, and their individual insight into the
questions posed within the study. Interviews were conducted in the local
language (Chichewa), audio recorded, and transcribed into English.
Participation was voluntary, and responses were recorded anonymously.
Written, informed consent was obtained from each respondent prior to the
interview. The study was approved by the National Committee on Research
in the Social Sciences and Humanities (NCRSH) of Malawi; Protocol NO.
P.03/19/356. Collected data were analysed thematically and was coded
within the software programme Nvivo, which organises materials and
assists with the coding process.

### Limitations

Though extensive and useful in understanding the conditions at QECH, the
data set would benefit from a comparison with background values. As
trash-burning is extensive throughout the city, there were no obvious
locations that could be confidently used as representative background
values. And although there were several possible areas upwind and
outside of town, the logistics and safety involved in accessing them to
change the batteries meant that the measured values at Queens could not
be compared to a stable background concentration. Knowing the wind
direction and velocity would have helped in better identifying the
potential sources of the measured PM. However, the necessary equipment
was not easily available in Blantyre at the time, and as an exploratory
study, the immediate hazards for the staff and patients at QECH are
relevant regardless of the burning source.

Knowing the exact burning locations and burning times, especially at the
incinerator and the Guardian Shelter, would have helped in better
identifying the sources and movements of the plumes. However, given the
size of the campus and it's 24-hour schedule, a much larger research
team would have been required to quantify all the burning. Finally, The
humidity, and the potential for interference was not accounted for; the
measurements were not adjusted for humidity which could affect the
values, though not enough to significantly alter the key findings and
need for immediate action.

# Theory/calculation (not applicable)

# Results and Discussion

The Malawi Bureau of Standards is the national agency responsible for
setting and publishing all standards in the country. At the time of
writing, the official website (www.mbsmw.org) was not available.
However, published work that references Malawian standards
[@kutlarjoss2017time; @mapoma2014air] indicate that the maximum 24-h
PM~10~ value is 25 µg/m^3^ and the maximum annual PM~2.5~ value is 8 µg/m^3^.
There is no daily maximum value for PM~2.5~. It should be noted that both
of these values are at, or below, the level-4 interim WHO targets.

Therefore, for international comparisons and due to the lack of
comprehensive Malawian Standards, we use the WHO recommendations as a
basis of comparison for the measured values
[@worldhealthorganization2021who]. The relevant particulate matter
targets are presented in @tbl-who-targets.

 (TODO: REPLACE screenshot)

```{r}
#| label: tbl-who-targets
#| tbl-cap: Particulate matter targets from [@worldhealthorganization2021who]

knitr::include_graphics(here::here("manuscript/img/image-replace-table.png"))

```

## Particulate Measurements

### Peaks

Due to localised variation (air movement in the immediate area) and the
density of the observations, the plotted 5-minute data obscure
persistent trends. The full set of 5-minute data are plotted and
presented in Appendix A @fig-five-minute-intervals. However, the number of measured values
that exceeded the hazardous limit for both parameters are summarised in
@tbl-peaks.

```{r}
#| label: tbl-peaks
#| tbl-cap: Peaks for both PM~2.5~ and PM~10~ are the number of data points above the WHO interim target 1 (annual) which is the least stringent; the number of observations recorded (count) is provided for reference.

observations_location <- dat_in |> 
  select(date_time, location) |> 
  unique() |> 
  count(location)

no_peaks_who <- dat_in |> 
  mutate(peaks = case_when(
    indicator == "pm25" & value >  35 ~ TRUE,
    indicator == "pm10" & value >  70 ~ TRUE,
    TRUE ~ FALSE
  )) |> 
  mutate(date = as_date(date_time)) |> 
  group_by(indicator, location) |> 
  summarise(
    peaks = sum(peaks)
  ) |> 
  ungroup() |> 
  pivot_wider(names_from = indicator, values_from = peaks)

observations_location |> 
  left_join(no_peaks_who) |> 
  knitr::kable()
```

Over the course of the two-month study, the hazardous limits for both
parameters were exceeded at all locations except for the Lions Sight
hospital. The limits were exceeded fewer than 50 times at five of the
locations, and only the monitors at the Lighthouse Clinic and the
Guardian Shelter recorded more than 50 instances above the hazardous
limits. At the Lighthouse Clinic, the PM~10~ limit was exceeded almost
twice as often as the PM~2.5~ limit, while at the Guardian Shelter, the
reverse is true. Because the Lighthouse Clinic is within 50 m of the
incinerator, the number and predominance of PM~10~ peaks is characteristic
of incomplete combustion and dust that is typical for the area and the
open burning that occurs there. The residents at the Guardian Shelter
however are mostly cooking within contained clay stoves that have been
designed to improve combustion of the coal and wood that is burned
within them. The fewer number of PM~10~ peaks is a testament to this
intervention, though the frequency of PM~2.5~ peaks is still beyond
acceptable.

### 24-hour averages

The 24-hour averaged values (logged every 5 minutes) are presented for
each location, for both PM~10~ and PM~2.5~ in fig-24-hour-average. As well as dampening
local variation and peaks, the WHO air quality guidelines are also based
on 24-hour averages which are the standard against which the health
risks can be judged.

Overall, PM~2.5~ values remained below 100 µg/m^3^ at 6 of 8 locations
(Administration, Lions Sight, Malaria Project, uMoyo, Ward 6B, and Mercy
James); PM~10~ values were consistently below 100µg/m^3^ at the same
locations, but with several average values extending slightly above, and
then infrequently.

The daily averages at both the Lighthouse Clinic and the Guardian
Shelter are both consistently higher for both parameters and the two
averages closely followed the same general trends. Though the peaks at
the Lighthouse Clinic were higher than the Guardian Shelter, the low
values were consistently lower, indicating more times of little or no
burning, unlike the Guardian Shelter emissions which were relatively
constant. However, further analysis of 12 hour averages (8:00-15:59
(working hours) and 16:00-7:59 (evening)) did not indicate clear
differences between the time periods; stated differently, the data did
not clearly point to more or less burning in the day or at night.

```{r}
#| label: fig-24-hour-average
#| fig-cap: Average 24-hour PM~10~ and PM~2.5~ at 8 monitoring stations over 2 months
#| fig-align: center

dat_in_sum_day |> 
  ggplot(aes(x = date, y = mean, group = indicator, color = indicator)) +
  geom_point(size = 0.8) +
  geom_line() +
  labs(x = NULL) +
  ylab(bquote(uq/m^3)) +
  facet_wrap(~location, ncol = 4) + 
  scale_y_continuous(limits = c(0, 400)) +
  scale_color_brewer(type = "qual", palette = 2) + 
  theme_bw(base_size = 14) + 
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())

```

### PM ratios

Given that the type of fuel and the type of burning (contained vs. open)
produce very different particulate "fingerprints", the ratio of PM~10~ to
PM~2.5~ values were examined to determine if a clear difference between
locations, and therefore source, could be identified. The results are
presented in @fig-pm-ratios.

```{r}
#| label: fig-pm-ratios
#| fig-cap: PM~10~:PM~2.5~ values for each location. Each line shows the minimum and maximum ratio for the day based on ratios calculated hourly. Does not include 14th to 16th October at location uMoyo due to extreme outliers.
#| fig-align: center

dat_in_ratios <- dat_in |> 
  mutate(date = as_date(date_time)) |> 
  mutate(hour = hour(date_time)) |> 
  group_by(date, hour, location, indicator) |> 
  summarise(mean = mean(value)) |> 
  pivot_wider(names_from = indicator, 
              values_from = mean) |> 
  mutate(p10_pm25_ratio = pm10/pm25)  |>

  filter((date != "2019-10-14" & date != "2019-10-15" & date != "2019-10-16")
         | location != "uMoyo") |> 
  #filter(p10_PM~2.5~_ratio > 30)

  group_by(date, location) |> 
  summarise(min = min(p10_pm25_ratio),
            max = max(p10_pm25_ratio))

# Plot

dat_in_ratios |> 
  ggplot(aes(x = date)) +
  geom_linerange(aes(ymin = min, ymax = max)) +
  labs(x = "hour") +
  facet_wrap(~location, ncol = 2)
```

The results presented in Figure 5 illustrate both the presence of PM~10~
relative to PM~2.5~ as well as the range of values over which that ratio
is observed.

Unlike the peak values presented in Table 2, the values presented in
Figure 5 indicate the relative presence of larger (PM~10~) particles to
smaller ones (PM~2.5~) regardless of their concentration: the ratio of two
low concentrations can be the same as the ratio of two high
concentrations and is therefore more indicative of source than of
proximity.

In general, the calculated ratios are concentrated between 1.25 with few
values exceeding 7.5. Three from each the Guardian Shelter and the Ward
6B exceeded 7.5 though the most values exceeding 5 were at the Guardian
Shelter. It is interesting to note that despite the large daily
variations at the Lighthouse Clinic, the parameter ratios there were
fairly consistent with few spikes, and only 2 days with ratios above 5.

An examination of the range of ratios is used to better understand the
variability within a single day. For example, several of the calculated
values at the Guardian Shelter have a range of more than 5, and one day
had a range of close to 10 i.e. the minimum ratio recorded for that day
was close to 1 and the maximum value was over 10. The composition of
emissions recorded at that location ranged significantly, and likely
reflects a range of burning styles and/or fuel type.

## Exposure Data

The relative amount of time exposed to a given category of air quality
is shown in @fig-percent-exposure. Each square represents 1% of the measured time.

```{r}
#| label: fig-percent-exposure
#| fig-cap: Percentage of measured values categorised according to WHO 2021 targets
#| fig-align: center

dat_air_quality_sum_who_2021 |> 
  #filter(indicator == "PM~2.5~") |> 
  group_by(location, indicator) |> 
  mutate(exposure_percent = round(exposure_hrs / sum(exposure_hrs) * 100, 0)) |>

  # manipulate data for purpose of plot

  mutate(exposure_percent = case_when(
    location == "Administration" 
    & indicator == "pm10"
    & air_quality_who_annual == "Interim target 2" ~ exposure_percent - 1,

    location == "Administration" 
    & indicator == "pm25"
    & air_quality_who_annual == "Interim target 3" ~ exposure_percent + 1,

    location == "uMoyo" 
    & indicator == "pm25"
    & air_quality_who_annual == "Interim target 4" ~ exposure_percent - 1,

    location == "Mercy James" 
    & indicator == "pm10"
    & air_quality_who_annual == "Interim target 4" ~ exposure_percent + 1,

    TRUE ~ exposure_percent
  )) |>

  #filter(is.na(exposure_percent))

ggplot(aes(fill = air_quality_who_annual, values = exposure_percent)) +
  geom_waffle(color ="white", size = 0.5, n_rows = 5, na.rm = TRUE) +
  coord_equal() +
  labs(fill = "WHO 2021 targets (annual)") +
  scale_fill_brewer(type = "div", palette = "RdYlBu", direction = -1) +
  scale_x_discrete(expand=c(0,0)) +
  scale_y_discrete(expand=c(0,0)) +
  facet_grid(location ~ indicator, switch = "y") +

  #facet_wrap(~location, ncol = 2) +
  theme_minimal(base_size = 14) +
  # help from: https://stackoverflow.com/questions/34749636/left-justify-text-from-multi-line-facet-labels
  theme(strip.text.x = element_text(hjust = 0),
        strip.text.y = element_text(vjust = 0, size = 4))
```

Unsurprisingly, a majority of the air quality at the Lighthouse Clinic
and the Guardian Shelter falls at, or below, interim target 1. What is
less expected is that even though there are fewer days of hazardous or
very bad air quality at the other locations, most locations do not have
target 4 values for even half of the time. Mercy James, the Malaria
Project and Lions Sight have the largest percentages of days with air
quality meeting the interim target 4. Interim target 2 rated air
occupies an unexpectedly large proportion of the days at all locations,
indicating that although there are a few very toxic, and a few quite
good days, the majority of the air is actually neither, but still cause
for concern.

## "The Air We Breathe is Not Good": Perspectives from QECH

As the previous section has described, the open incineration of waste at
QECH has created hazardous conditions for those occupying the space.
These risks were not lost on staff and caregivers, as interviews
demonstrated broad and near universal awareness of the poor air quality
within the hospital grounds. Overwhelmingly, within interviews, both
caregivers and staff were quick to decry the poor quality of the air,
generally without prompting. The few exceptions were those staff posted
on the peripheries of the hospital grounds, at a distance from the
spaces of incineration. However, even those who did not experience the
impacts of the burning, were aware of it, and considered themselves
fortunate to be posted in a section of the hospital where it was less of
a problem. Furthermore, according to caregivers and staff who work night
shifts, air quality can be particularly bad at night and in the early
morning, despite the smoke being less visible, due to the habit of
janitorial staff concentrating their burning during the late hours. The
poor air quality on hospital grounds is also a frequent cause for
complaint by patients and visitors, with nearly every staff member
interviewed being able to recall having received a complaint, and in
turn, complaining to administration. One of the staff members
responsible for the burning said that he personally, had received
hundreds of complaints, but was powerless to affect meaningful change,
aside from burning at different hours, until the incinerator could be
repaired.

Nonetheless, despite this consensus that air quality caused by the
burning of waste within the hospital was a problem, significant
differences emerged between respondents over their understandings of
potential impacts, the effectiveness of various coping mechanisms, and
their problematisations linked to the burning of specific waste
materials. In addition, interviews revealed that these understandings
were informed by a significant amount of misinformation, even amongst
trained medical staff, which may contribute to them being less able to
mitigate potential risks for themselves and those who rely on their
care.

### Impacts, Problematisations, and Misinformation

The poor air quality within QECH was responsible, according to
respondents, for a wide array of health impacts. The most common ones
cited included: coughing and sneezing, sore throat, stinging eyes,
breathing difficulties, and persistent cold and flu. Nausea was also
mentioned, but was not a commonly described impact. Only one respondent,
of the 26 total interviewed, did not describe lingering health impacts
which they could ascribe to the smoke, however, they also described
having chronic eye irritation, but did not believe the smoke was a
contributing factor.

In addition to these impacts, which respondents bear on a daily basis,
many also believed that the smoke could contribute to a number of more
serious, long-term health complications. For instance, nearly a quarter
of respondents raised concerns of the potential impact that the smoke
could have on patients or staff with asthma. Others flagged poor air
quality as a potential risk factor for certain cancers, lung disease, or
heart problems. For a few, the smoke posed an unknown danger, they were
not sure what types of impacts it could have, but they were sure it was
harmful in some way.

Also, understandably, given the large tuberculosis ward present within
the hospital grounds, and the high prevalence of the disease within
Malawi [@worldhealthorganization2019global] there was significant
concern (more than half of respondents) about the impact the poor air
quality could have on the infected. However, there also persisted a
belief among several respondents, including several nurses, that the
smoke could be a cause of the disease itself. As one staff member
(20/11/2019) stated, "I believe breathing this air for a long time can
cause Tuberculosis." This, however, was only one of the few instances of
misinformation which staff members held regarding air quality and
health. Another example, voiced by several respondents, included a
belief that some staff members were immune to ill-health impacts of the
smoke, because they had received vaccinations from the hospital.
Nonetheless, they were concerned about the impacts of the smoke on
patients and visitors, as one staff member (18/11/2019) expressed:

> Personally, I have never experienced \[eye discomfort\] because I get
> vaccinated and I am protected including other staff. However, we
> realize that the air can badly affect other people and patients who
> come to this place.

Another interesting misconception that emerged, which may be tied partly
to translation and transcription, was a different cultural understanding
of smoke versus smells. More than half of the respondents appeared to
conflate the two, with some expressing a belief that it was the odour of
what was being burnt that was harmful, rather than the smoke being given
off. This has led many to specifically problematise the burning of
certain wastes, such as plastics, medicines, and other medical wastes,
which give off distinctive or less pleasant odours, as opposed to the
burning of other items, such as garden refuse, which may produce
significant smoke, and contribute to higher recorded values of
particulate matter, but produce a less pungent, or more normalised,
odour.

### Coping Mechanisms

Finally, in order to manage the impacts of QECH's persistently poor air
quality, staff and visitors reported having developed a number of coping
mechanisms, designed to help them get through their daily routines.
These included staying indoors, blocking doors and windows, and taking
breaks away from hospital grounds in order to catch some breaths of
fresher air. However, for janitorial staff, inside was not necessarily
better, as several reported their indoor workspaces as being also
insufferable for long periods of time from the smell of cleaning agents
and other chemicals. Furthermore, amongst respondents there was a
general disagreement over the effectiveness of personal protective
equipment (PPE), such as face masks, towards mitigating the impacts of
the smoke. A few staff described pleading to hospital administration for
such equipment, to no avail. However, other staff members, who do have
access to PPE, noted that even face masks do little to mitigate the
impacts of the smoke, describing them as ineffective.

Most staff, however, have been unable to find anyway to mitigate the
impacts of the smoke, and only found relief once they reached home at
the end of their shift, as one of the janitorial staff described, "we
only feel safe when we are home". Of course, this relief is not
available for the hundreds of patients and caregivers who sleep at QECH
or are unable to leave. Ultimately, most place their hope in the
construction of the new incinerator (which had not yet been activated at
the time of the interviews), and biding their time as construction drags
on; coping as best they can. As one staff member (25/11/2019)
described,"we are just hoping we will start breathing good air soon,
when the new incinerator is opened."

# Conclusions

The air quality, as a result of open waste burning at Queen’s is poor and not suitable for a city, let alone a hospital. Over the course of this two-month study, the hazardous limits for both PM parameters were exceeded at all locations except for the Lions Sight hospital. The WHO limits were exceeded fewer than 50 times at five of the locations, and only the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the most hazardous limit. The composition of emissions recorded ranged significantly, analysed as the ratio PM~10~:PM~2.5~ likely reflects a range of burning styles and/or fuel (waste) type, and no discernable trend was observed..  Among interview respondents, there was a general  consensus that air quality caused by the burning of waste within the hospital was a problem, although there were significant differences between respondents over their understandings of potential impacts, the effectiveness of various coping mechanisms, and their problematisations linked to the burning of specific waste materials. For most, going home or leaving the Queen’s campus was the only pathway to relief, though for patients who can not leave, endurance was the only option. This work was originally planned as a before-and-after study:  we envisioned reporting these baseline results compared to follow-up measurements once the long-promised incinerator was commissioned.  However, nearly 4 years later, the incinerator is rarely functional (partly for technical reasons, partly for financial ones) and the smoke hovering over Queen’s is still thick.  Waste management has long been known as an environmental health issue, but in the case of open trash burning at a medical facility, it becomes more:  for some of the most vulnerable, especially those with TB, asthma or other chronic disease, it could be a matter of life or death.

# Acknowledgements

The authors are grateful to the staff, patients and caregivers at QECH for their assistance, willingness to share their views, and importantly, perseverance in delivering the best care possible under difficult circumstances. Moreover, we would like to acknowledge the assistance and effort of Open Data Durban, now the Open Cities Lab, for their assistance in constructing and programming the sensors.

# Appendix {.unnumbered}

If there is more than one appendix, they should be identified as A, B,
etc. Formulae and equations in appendices should be given separate
numbering: Eq. (A.1), Eq. (A.2), etc.; in a subsequent appendix, Eq.
(B.1) and so on. Similarly for tables and figures: Table A.1; Fig. A.1,
etc.

## Appendix A {.unnumbered}

```{r}
#| label: fig-five-minute-intervals
#| fig-cap: PM~2.5~ and PM~10~ values collected every 5 minutes over 10 months at 8 locations.
#| fig-align: center

dat_in |> 
  ggplot(aes(x = date_time, y = value)) +
  geom_line(alpha = 0.5) +
  labs(title = "Hospital Air Quality",
       subtitle = "PM~2.5~ and PM~10~ measurements in 5-min intervals at eight locations") +
  ylab(bquote(uq/m^3)) +
  facet_grid(indicator~location) +
  theme_bw(base_size = 14) + 
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

## Appendix B {.unnumbered}

### Hardware Components {.unnumbered}

1.  Raspberry Pi 3 Model B single-board computer technical
specifications:
-   Broadcom BCM2837 64bit Quad Core Processor powered Single Board
Computer running at 1.2GHz 1GB RAM
-   BCM43438 WiFi on board
-   Bluetooth Low Energy (BLE) on board
-   40pin extended GPIO
-   4 x USB 2 ports 4 pole Stereo output and Composite video port
-   Full size HDMI CSI camera port for connecting the Raspberry Pi
camera
-   DSI display port for connecting the Raspberry Pi touch screen
display
-   Micro SD port for loading your operating system and storing data
-   Upgraded switched Micro USB power source (now supports up to 2.4
Amps)
-   Same form factor as the Pi 2 Model B, however the LEDs have
changed position
-   RPI3-MODB-16GB-NOOBS
2.  Nova PM SDS011 Sensor
-   See Data Sheet attachment for technical specifications
3.  MCP3008 microchip
-   analog-to-digital converter (ADC)
4.  Romoss Sense8+ 30000mAh QC Type-C
-   Power bank battery
5.  Pipe
-   clear rubber pipe attainable from any local hardware store
6.  Box
-   any box that can fit all the components comfortbaly can be used.
The box will need to have a whole drilled in it for the pipe.

### Software ‘components’ and data management {.unnumbered}

1. The Raspberry Pi operating system runs off of Rasbian software on a NOOBS scandisk. 
2. Air quality monitoring code is written in Python using the [aqi library](https://pypi.org/project/python-aqi/) (A library of algorithms to convert between AQI value and pollutant concentration). The code is stored on the [Open Data Durban GitHub account](https://github.com/opendatadurban/hospital_stations) and is free for all to access. To be able to collect data, the device needs to be connected to a reliable wifi network, the SSID and password need to be coded onto the device during step 2 in the building method.
3. Once the whole device has been put together and placed on site, to access that data we use a combination of propgrams:
- [Putty](https://www.putty.org/) - this is an SSH client that we use to gain access to the Raspberry Pi from another computer. To be able to access the Raspberry Pi in question, the computer in use and the air quality monitoring device need to be on the same network for the connection to be successful. 
- [Winscp](https://winscp.net/eng/download.php) - this is an FTP client that we use, in conjunction with Putty, to be able to transfer data files from the air quality monitoring device to one’s computer.
- [Remote.it](https://app.remote.it/) - this is an application that supports remote SSH, so this is used when wanting to access data from a device that is not connected to the same network as the computer in use. Remote.it is used together with Putty and Winscp to access the air quality data remotely.

### Connection/building method {.unnumbered}

To build the sensor, we followed these steps:

1. Set up the Raspberry Pi with the Rasbian software using the NOOBS scandisk. (insert the scandisk into the Rasberry Pi) A comprehensive start up guide can be found here. 
2. Next we load the code onto the Raspberry Pi. The code can be found on the Open Data Durban Hospital Stations Repository Github. To use it you can follow the steps in the file called Initiate_WS.txt. Once all the steps have been completed, the Raspberry Pi is now a device that can be connected to a particulate matter monitoring sensor.  
3. Connect the hardware components
- Connect the Raspberry Pi to the Nova PM Sensor using the break out board that comes with the Sensor. 
- Connect the battery to the Raspberry Pi via the micro port
- Attach the rubber pipe to the sensor 
- Assemble all components inside the box, with the pipe coming out of the drilled hole
- To prevent dirt entering the system, close off all openings between the pipe and the box with silicone putty
- Secure box in an elevated position (not too close to the ground) and with the pipe facing the direction of the inflow of the air under investigation.

# References {.unnumbered}