### 2023-04-18 README.qmd
Re-organise content

---
format: gfm
execute: 
  echo: false
  warning: false
  freeze: true
editor: 
  markdown: 
    wrap: 72
bibliography: [manuscript/man-qechairquality.bib, manuscript/packages.bib]
---

## Manuscript: "'It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi"

This repository contains a reproducible manuscript for submission to the
journal [Habitat
International](https://www.sciencedirect.com/journal/habitat-international).

## Data

Raw data and analysis-ready processed data is available as an R package
at: https://github.com/Global-Health-Engineering/qechairquality [@R-qechairquality].

The data underlying the tables and figures of this manuscript are
contained in this repo in [data/processed-data/](https://github.com/Global-Health-Engineering/manuscript-qechairquality/tree/main/data/processed-data).

## Figures

Figures were prepared using the [Elsevier guidelines for artwork and
media](https://www.elsevier.com/authors/policies-and-guidelines/artwork-and-media-instructions).
The exported figures in TIFF format exceeded 100 MB per file and are not
tracked in this repository. The automated exported figures from the
Quarto markdown file [are accessible in the
repository](https://github.com/Global-Health-Engineering/manuscript-qechairquality/tree/main/manuscript/manuscript_files/figure-pdf).

## Software

R Statistical Software version 4.2.1, RStudio IDE version 2023.3.0.386,
and Quarto scientific publishing system version were used for
quantitative data analysis and writing of the manuscript [@R-base;
@allaire2022quartoa; @positteam2023rstudio].

A set of additional R packages were used for data wrangling, analysis,
and visualisation [@ggplot22016; @lubridate2011; @R-dplyr; @R-forcats;
@R-ggplot2; @R-janitor; @R-leaflet; @R-lubridate; @R-mapview; @R-readr;
@R-tidyr; @R-waffle].

The [Quarto template for Elsevier journals](https://github.com/quarto-journals/elsevier) was used to prepared the final
submitted PDF document.

## References