### 2023-05-19 manuscript/manuscript-qech-air-quality-tilley-credit.qmd
Submitted manuscript

---
title: "'It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi"
author:
  - name: Elizabeth Tilley
    orcid: 0000-0002-2095-9724
    email: tilleye@ethz.ch
    affiliations: 
        - id: ethz
          name: ETH Zurich
          department: Department of Mechanical and Process Engineering
          address: Global Health Engineering, Clausiusstrasse 37
          city: Zurich
          country: Switzerland
          postal-code: 8092
    attributes:
        corresponding: true
  - name: Hope Chilunga
    orcid: 0000-0003-4216-6319
    email: hchilunga@poly.ac.mw
    affiliations: 
        - id: mubas
          name: Malawi University of Business and Applied Sciences
          department: Department of Environmental Health
          city: Blantyre
          country: Malawi
  - name: Jonathan Kwangulero
    orcid: 0000-0002-4050-5121
    email: kwangulero@gmail.com 
    affilations:
        - ref: mubas
  - name: Lars Schöbitz
    orcid: 0000-0003-2196-5015
    email: lschoebitz@ethz.ch
    affiliations:
        - ref: ethz
  - name: Saloni Vijay
    orcid: 0000-0001-5883-5308
    email: svijay@ethz.ch
    affiliations:
        - ref: ethz
  - name: Heiko Heilgendorff
    orcid: 0009-0004-1389-5541
    email: heiko.heilgendorff@gmail.com
    affiliations: 
        - id: ukzn
          name: University of KwaZulu-Natal
          department: School of Mathematics, Statistics and Computer Science
          city: Durban
          country: South Africa
  - name: Marc Kalina
    orcid: 0000-0002-6335-3845
    email: mkalina@ethz.ch
    affiliations:
        - ref: ethz
date: last-modified
bibliography: [man-qechairquality.bib, packages.bib]
format:
  docx: default
execute: 
  echo: false
  warning: false
  freeze: true
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: console
---

# CRediT roles

```{r}

library(tidyverse)

df_credit <- read.csv("https://raw.githubusercontent.com/Global-Health-Engineering/manuscript-qechairquality/main/data/credit.csv") |> 
  unite("name", first_name:surname, sep = "")

```

```{r}
credit_sentence <- larnsceR::credit_sentence(data = df_credit, group_var = name)
```

`r credit_sentence$sentence`